---
title: "Intro into R"
author: "vitaly.belik@fu-berlin.de"
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output:
  html_document:
    depth: 3
    fig_caption: yes
    highlight: tango
    number_sections: yes
    theme: paper
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Basics
## Show the current directory
```{r}
getwd()
```
## Creating arrays of values
```{r}
y <- c(1,10)
y
u <- c(1:10)
u
v <- u^2
v
```
## Creating dataframe
```{r}
df <- data.frame(u,v)
df
```
## Writing dataframe to file
```{r}
write.csv(df, file = "MyData.csv")
list.files()
```
## Reading data from file
```{r}
ndf <- read.csv("MyData.csv")
colnames(ndf)
summary(ndf)
```
## Deleting a column
```{r}
Data <- subset( ndf, select = -X )
Data
```
## Adding a column
```{r}
Data['w'] <-u
Data
```
## For-loop and if-condition
```{r}
for (i in c(1:10)){
  print(c("index :",i))
  if ((i > 3) & (i<10)){
    print("within if-condition")
    }
}
```
## Subsetting / filtering data
```{r}
Data[Data$u == 1,]
Data[Data$u == 2,]
?subset
subset(Data, u > 1)
```
## Creating plots
```{r}
plot(u,v)
```

# Working with real Data and using grammar of graphics (ggplot)
```{r}
library(ggplot2)
```


## Lactation curve
```{r}
#d = read.table('ReadData.csv', header = TRUE, sep = ";", dec = ',')
d = read.table('RealData.csv', header = TRUE, sep = ',')
ggplot(d, aes(x=DIM, y=GEMELK)) + geom_point() 
```
## Lactation curve
```{r}
#d = read.table('ReadData.csv', header = TRUE, sep = ";", dec = ',')
d = read.table('RealData.csv', header = TRUE, sep = ',')
ggplot(d, aes(x=DIM, y=GEMELK, colour = factor(Kuh))) +
  theme(text = element_text(size=18)) +
  geom_point(alpha = 0.3) +
  labs(x = "cc", y = "milk")
```

## Box plots

```{r}
#ggplot(data = d, aes(x = LAKTATION, y = GEMELK)) + geom_bar()
p <- ggplot(data = d, aes(x=LAKTATION, y=GEMELK, fill=factor(LAKTATION)))
p + geom_boxplot()
```

## Box plot with jittered points

```{r}
# 0.2 : degree of jitter in x direction
p + geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2))
```

# Violin plots

```{r}
p + geom_violin()
```


## Values distributions

```{r}
ggplot(d, aes( x = GEMELK, colour = factor(LAKTATION))) + geom_density()
```

## Curve fitting

```{r}
 ggplot( d, aes( x = DIM, y = GEMELK )) + geom_point( aes( colour = factor(LAKTATION)), shape = 19 ) + geom_smooth( method = "loess" ) + scale_colour_discrete( guide = FALSE )
```

```{r}
ggplot( d, aes( x = DIM, y = GEMELK )) + geom_point( aes( colour = factor(LAKTATION)), shape = 19 ) + scale_colour_discrete( guide = FALSE )+geom_density2d()
```
